kNN Implementation

### 2020-10-05 DATA622/M05-kNN.rmd
Updates to KNN files

```{r}
euclideanDist <- function(a, b){
  d = 0
  for(i in c(1:(length(a)) ))
  {
    d = d + (a[[i]]-b[[i]])^2
  }
  d = sqrt(d)
  return(d)
}
```
Let us write another function to iterate over the test set 
and over the training set for each observation in the test set.
```{r}
knn_predict2 <- function(test_data, train_data, k_value, labelcol){
  print(k_value)
  print(labelcol)
  pred <- c()  #empty pred vector 
  #LOOP-1
  for(i in c(1:nrow(test_data))){   #looping over each record of test data
    eu_dist =c()          #eu_dist & eu_char empty  vector
    eu_char = c()
    good = 0              #good & bad variable initialization with 0 value
    bad = 0

@@

    head(eu[1:k_value,])
    tbl.sm.df<-table(eu$eu_char)
    cl_label<-  names(tbl.sm.df)[[as.integer(which.max(tbl.sm.df))]]

    pred <- c(pred, cl_label)
  }
  return(pred) #return pred vector
}

```
Let us also write a utility function for accuracy.
```{r}
accuracy <- function(test_data,labelcol,predcol){
  correct = 0
  for(i in c(1:nrow(test_data))){
    if(test_data[i,labelcol] == test_data[i,predcol]){ 
      correct = correct+1
    }
  }
  accu = (correct/nrow(test_data)) * 100  
  return(accu)
}
```
Now let us load and prepare training/test set from the heart dataset.

@@

```
Computing ROCR and AUC is some what non-trivial as kNN do not compute probabilities
and results in unreliable ROCR plots.
```{r}
test.df[,predictioncol] <- predictions #
print(paste('accuracy=',accuracy(test.df,labelcol,predictioncol)))
```
Now let us label the test set using standard kNN implementation from class
package.
```{r}
library(class)
library(ROCR)

>>>>

```{r}
euclideanDist <- function(a, b){
  d = 0
  for(i in c(1:(length(a)) ))
  {
    d = d + (a[[i]]-b[[i]])^2
  }
  d = sqrt(d)
  return(d)
}
```
Let us write another function to iterate over the test set 
and over the training set for each observation in the test set.
```{r}
knn_predict2 <- function(test_data, train_data, k_value, labelcol){
  #print(k_value)
  #print(labelcol)
  pred <- c()  #empty pred vector 
  #LOOP-1
  for(i in c(1:nrow(test_data))){   #looping over each record of test data
    eu_dist =c()          #eu_dist & eu_char empty  vector
    eu_char = c()
    good = 0              #good & bad variable initialization with 0 value
    bad = 0

@@

    #head(eu[1:k_value,])
    tbl.sm.df<-table(eu$eu_char)
    cl_label<-  names(tbl.sm.df)[[as.integer(which.max(tbl.sm.df))]]

    pred <- c(pred, cl_label)
  }
  return(pred) #return pred vector
}
```
Let us also write a utility function for accuracy.
```{r}
accuracy <- function(test_data,labelcol,predcol){
  correct = 0
  for(i in c(1:nrow(test_data))){
    if(test_data[i,labelcol] == test_data[i,predcol]){ 
      correct = correct+1
    }
  }
  accu = (correct/nrow(test_data)) * 100  
  return(accu)
}
```
Now let us load and prepare training/test set from the heart dataset.

@@

```

```{r}
knn_predict3 <- function(test_data, train_data, k_value, labelcol){
  pred <- c() # Initialize an empty vector to hold the predictions
  eu_char <- train_data[,labelcol] # Get the class labels from the training data
  b <- train_data[,-c(labelcol)] # Drop the class label from the training data
  for(i in c(1:nrow(test_data))){
    a <- test_data[i,-c(labelcol)] # Pull the row from the test data
    eu_dist <- euclideanDist(a, b) # Calculate the euclidean distance between training data and test row
    eu <- data.frame(eu_char, eu_dist) # Put the distance and the label in a data frame
    eu <- eu[order(eu$eu_dist),] # Sort it by the euclidean distance
    eu <- eu[1:k_value,] # Subset the first k rows
    tbl.sm.df<-table(eu$eu_char) # Create frequency table
    pred <-  c(pred, names(tbl.sm.df)[[as.integer(which.max(tbl.sm.df))]]) # Add in the predicted class (highest frequency) to the predictions vector
  }
  return(pred) # Return the predictions
}
predictions2 <- knn_predict3(test.df, train.df, K,labelcol)
```

```{r}
start_time <- Sys.time()
n_tests <- 100
for (i in 1:n_tests){
  knn_predict2(test.df, train.df, K,labelcol)
}
end_time <- Sys.time()
difftime(end_time, start_time)
```

```{r}
start_time <- Sys.time()
for (i in 1:n_tests){
  knn_predict3(test.df, train.df, K,labelcol)
}
end_time <- Sys.time()
difftime(end_time, start_time)
```
Computing ROCR and AUC is some what non-trivial as kNN do not compute probabilities
and results in unreliable ROCR plots.
```{r}
test.df[,predictioncol] <- predictions #
print(paste('accuracy=',accuracy(test.df,labelcol,predictioncol)))
```
Now let us label the test set using standard kNN implementation from class
package.
```{r}
library(class)
library(ROCR)
